---
title: "Crime Rates"
---

Follow along with the demo to learn how to calculate crime rates using R and the Tidyverse. We'll also introduce a new library, janitor, which helps make column names easier to work with. You should get started by loading the tidyverse in a new R block, then install and load janitor.

### Task 1: Load the tidyverse and install and load janitor
```{r}
library(tidyverse)
install.packages("janitor")
library(janitor)
```


### Task 2: Load the crime dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv

```{r}
md_crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv") %>% clean_names()
```


### Task 3: Use janitor to clean up the column names

Done in task 2^^

### Task 4: Calculate the total crime rate for each year (picking our unit!)

```{r}
md_crimes_tot <- md_crimes |> mutate (total_rate = grand_total/population*10000)
```

### Task 5: Calculate the violent and property crime rates for each year (same unit)

```{r}
md_crimes_viol <- md_crimes_tot |> mutate (violent_rate = violent_crime_total/population*10000)
final_md_crimes <- md_crimes_viol |> mutate (property_rate = property_crime_total/population*10000)
```


### Task 6: Describe our results

Write a few sentences about what you found in the data. What years had the highest crime rates? What years had the lowest crime rates? Does the pattern match your expectations?

Throughout the 2000s, the crime rates seem to be consistently decreasing, with 2020 having the lowest crime rate overall. Despite this, violent crimes have stayed relatively the same and the drop seems to be mostly from property crimes. Even so, the lowest violent crime rates have been in the 2000s and the highest in the '90s. The highest overall crime rates, landing in the 600s, were mostly in the '80s and '90s. I'm not quite sure what I expected to be honest but I am definitely glad that at least some crimes have decreased!

When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

