---
title: "Volleyball Percentiles"
---

We'll be calculating percentiles to see how good Maryland was in service aces in the 2024 season. We'll use `mutate()` and a new function `ntile()` to calculate percentiles. Get started by loading the tidyverse in a code block, then load the CSV file here: https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv" and save it to a dataframe called `matches`.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the data

```{r}
matches <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv")
```

### Task 3: Create totals of aces for each team using `group_by()` and `summarize()`

```{r}
hometeam_aces <- matches |>
  group_by(team) |>
  summarize(hometeam_aces = sum(aces)) |>
  arrange(desc(hometeam_aces))
hometeam_acesagainst <- matches |>
  group_by(team) |>
  summarize(hometeam_acesagainst = sum(defensive_aces)) |>
  arrange(desc(hometeam_acesagainst))
```

### Task 4: Calculate percentiles using `mutate()` and `ntile()`

```{r}
hometeam_aces |> mutate(percentile = ntile(hometeam_aces, 100))
hometeam_acesagainst |> mutate(percentile = ntile(hometeam_acesagainst, 100))
```

### Task 5: Describe the percentiles

Write a couple of sentences that puts Maryland into context with the other teams in the dataset. You can do some Internet research to help with this.

Maryland Volleyball played better offensively at their home games than at their away games. At home, the team hit 223 total aces, making it in the 95th percentile of all college volleyball teams this season. Away, the team hit 195 aces, which landed our team in the  significantly lower 81st percentile.

When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.



