---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

ANSWER HERE

There are some categories with significantly higher numbers (Accident, Accident with Impound, Auto Stolen, Theft, and Theft from Auto). A lot of these are car related which may be important to note. From a reporting standpoint, the homicides are typically more important since they're more severe, but so are the high numbers since it may show a trend. There doesn't seem to be too much of a trend of any one category or any significant spikes in the data.

### Task 4: Calculating Averages

Calculate the total number of incidents per month using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
total_incidents <- pg_crime %>% mutate(total = Jan+Feb+Mar+Apr+May+Jun+Jul+Aug+Sep+Oct+Nov+Dec)
average_incidents <- total_incidents %>% mutate(average = (total/12))
pg_crime_averages <- average_incidents
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

ANSWER HERE

The top four categories for crime in PG County are car related so I guess I'd be interested in why criminals are so interested in cars. Do people in PG County own cars that are easier to break into? Do they keep more valuables in their car? Do they own more expensive cars? Does the high number of accidents prove that Maryland drivers are actually as bad as people say they are?

The categories with the lowest are some of the more serious ones like homicide and sex offenses. I would ask what the police department is doing to handle these that the numbers are so low, what the original numbers have been over the years to see if there is a downward trend. There are a lot more questions I could ask but these are just the ones I'll put down for now.

### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
june_change <- pg_crime_averages %>% mutate((Jun-average)/Jun*100)
july_change <- june_change %>% mutate((Jul-average)/Jul*100)
august_change <- july_change %>% mutate((Aug-average)/Aug*100)
summer_crime_averages <- august_change

#Had a hard tme figuring out the right formula for this...
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

ANSWER HERE

Other Breaking and Entering crimes were way down, homicides were up. Different types of robberies were also down. Assault shootings were also up 25% in July and August. A lot of the crimes fluctuated a lot throughout the summer with a lot of them actually being below average. The theory is likely not true because a lot of the incidents are below average.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE

Prince George's county boasts high numbers of car accidents. 

Maryland has been the butt of many jokes concerning poor drivers, but there may be some truth to it. Out of 26737 crimes committed in 2024, almost 10,000 of them were car accidents. That's more than 1/3 of all crimes in the county. Car accidents pose a potentially fatal risk to both Maryland drivers and pedestrians. The more accidents that occur, the higher that risk is. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
