---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

ANSWER HERE

The spreadsheet lays out the name, status, title and salary of all White House employees. There are 11 employees on the list who have 0 listed under their salaries, which may be somehting ot look into and may skew some of the data calculations. There is also the difference between a detailee and an employee which may differentiate how you describe the data. Since detailees are typically paid by their original agency, these employees may not actually be being paid by the White House which may also skew the data.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries %>% 
  summarise(mean_wh_salaries = mean(salary))
wh_salaries %>% 
  summarise(median_wh_salaries = median(salary))
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries %>% 
  summarise(mean_wh_salaries = mean(salary))
wh_salaries %>% 
  summarise(median_wh_salaries = median(salary))
wh_salaries %>% 
  summarise(min_wh_salaries = min(salary))
wh_salaries %>% 
  summarise(max_wh_salaries = max(salary))
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

ANSWER HERE

#The median means that the middle number of the dataset is $110,500. The mean is $107036.6	which means that is the average. This number may be more affected by outliers. Without knowing the first and third quartile it's a bit difficult to calculate the IQR and determine if there are any outliers, but I think it's safe to assume that all of the zeroes are outliers, including the minimum, and so is the maximum.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

ANSWER HERE

The story idea I have that stands out the most is that the highest paid employee, is an associate counsel detailee who makes almost 60,000 more annually than the next highest paid employee for Biden's White House. Most associate counsel (n=8) for the Biden White House made $110,500 per year, which is the median salary for all White House employees. So why is Michelle M. Barrans making more than twice as much as many of the other associate counsel?

I would start off my looking at their workload and payment schedules if possible. Also, a lot of the detailees seem to be paid higher salaries than a lot of the other White House employees, so I may do some research on that. To be more specific with Ms. Barrans, I would look at the work she does compared to the other associate counsel to see if she is doing more work than her peers or any other factors that would explain this phenomenon.
