---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
songs <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")
```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
songs %>%  summarize(mean = mean(danceability), sd = sd(danceability))
```

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
songs %>% 
  ggplot() +
  geom_histogram(aes(x = danceability), binwidth = 0.01)
```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
songs %>% 
  ggplot() +
  geom_histogram(aes(x = danceability), binwidth = 0.01) +
  geom_vline(aes(xintercept = mean(danceability)), color = "purple", linetype = "dashed", linewidth = 0.5)
```


### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
songs %>% 
  ggplot() +
  geom_histogram(aes(x = danceability), binwidth = 0.01) +
  geom_vline(aes(xintercept = mean(danceability)), color = "purple", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) - sd(danceability)), color = "hotpink", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) + sd(danceability)), color = "hotpink", linetype = "dashed", size = 1)
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.

The distribution is very much skewed to the right, meaning more of the music is danceable than not. The mean is 0.673, which tells us that, on average, songs are pretty danceable. The fact that most of the songs fall within one standard distribution of the mean shows that most songs people were listening to in 2020 were pretty danceable.

Pitch: In 2020, Spotify users chose to listen to more danceable songs. Data from Spotify that ranks the "danceability" of over 1500 songs on the platform showed that users were listening to songs with an average danceability score of 67%.

Many of these songs came from popular artists like The Weeknd, Mac Miller, and Bad Bunny which doubles down on the current popularity of music that gets you on your feet.

### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
