---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
mnms <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10.
```{r}
sample5 <- mnms %>% 
  sample_n(5)
sample10 <- mnms %>% 
  sample_n(10)
```


### Task 4: Calculate the mean and standard deviation of the samples, using one of the colors.
```{r}
sample5 %>%  summarise(mean = mean(red), sd = sd(red))
sample10 %>%  summarise(mean = mean(red), sd = sd(red))
```


### Task 5: Calculate the mean and standard deviation of the population, using the same color as Task 4
```{r}
mnms %>%  summarise(mean = mean(red), sd = sd(red))
```


### Task 6: Compare the sample and population statistics

Compare the three sets of means and standard deviations. Are they similar?

The means and standard deviations from the two samples are somewhat close to each other but not too close. Both the means and the standard deviations of the samples are very far from the entire population's mean and standard deviation.

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

In order to sample randomly and accurately, you have to get a large enough proportion of the population or it is not a truly representative sample. The samples were too small to accurately analyze the entire population's data.The smaller the sample, the more inaccurate your results become.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
